Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-17 paper/paper(1).qmd
Add files via upload

```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE}
step_cons <- glm(vote_conservative ~ age+province+education+religion+income_before_tax,
          data=cleaned_survey_data, family = "binomial")
step_cons_counts <- cleaned_survey_data %>%
  group_by(age, province, education, religion, income_before_tax)%>%
  summarise(N_j=n(), .groups = 'keep')

@@

Party_Name <- c("Conservative", "Liberal")
yPS_primary <- c(cons_estimate,lib_estimate)
yPS_AIC <- c(cons_aic ,lib_aic)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(broom)

# Assuming you have already fitted a model `step_lib`
# and `step_lib` is an object of class 'glm' or similar
tidy_lib <- broom::tidy(step_lib, conf.int = TRUE, exponentiate = FALSE)  # Set exponentiate=TRUE for logistic regression

# Forest plot of coefficient estimates
ggplot(tidy_lib, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2) +
  geom_vline(xintercept = 1, linetype = "dotted", color = "red") +  # Use 0 if exponentiate=FALSE
  labs(x = "Estimate", y = "Coefficient", title = "Coefficient Estimates for Liberal Party") +
  theme_minimal()
# Assuming your `cleaned_survey_data` has a column `province` and `vote_liberal`
proportions_by_province <- cleaned_survey_data %>%
  group_by(province) %>%
  summarise(estimate = mean(vote_liberal)) %>%
  mutate(province = factor(province, levels = rev(unique(province))))  # Ordering the provinces

# Horizontal bar plot of estimated proportions by province
ggplot(proportions_by_province, aes(x = province, y = estimate)) +
  geom_col() +
  coord_flip() +  # Flipping the coordinates for horizontal bars
  labs(y = "Estimated Proportion", x = "Province", title = "Estimated Proportion of Voting for Liberal by Province") +
  theme_minimal()

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
cleaned_survey_data <- read_parquet(here("data", "analysis_data","cleaned_survey_data.parquet"))

cleaned_survey_data$vote_liberal <- ifelse(cleaned_survey_data$vote_liberal == "Yes", 1, 0)

lib_props <- cleaned_survey_data %>%

  group_by(age, sex, province, education, religion, income_before_tax) %>% 
  summarise(n = n(), .groups ='drop') %>%

  group_by(province) %>% 
  mutate(prop = n/sum(n))

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)
# Assuming 'cleaned_survey_data' is your data frame and it's already loaded in R
# First, check that all required variables are present in 'cleaned_survey_data'
required_vars <- c("vote_conservative", "age", "sex", "province", "education", "religion", "income_before_tax")
missing_vars <- setdiff(required_vars, names(cleaned_survey_data))

if(length(missing_vars) > 0) {
  stop(paste("The following required variables are missing in the data frame:", paste(missing_vars, collapse=", ")))
}

# If all variables are present, fit the linear model
model <- lm(vote_conservative ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)

# Use broom::tidy to get a tidy data frame of the model summary with confidence intervals
library(broom)
model_summary <- broom::tidy(model, conf.int = TRUE)

ggplot(model_summary, aes(x = estimate, y = reorder(term, estimate))) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Estimate", y = "Coefficient", title = "Coefficient Estimates") +
  theme_minimal()

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

library(ggplot2)
library(dplyr)

# Assuming 'your_data' is your dataframe and it contains the variables 'province' and 'vote_conservative'
# where 'vote_conservative' is a binary (0/1) indicating whether a vote for conservative was cast.

# Calculate proportions and confidence intervals by province
conservative_votes_by_province <- cleaned_survey_data %>%
  group_by(province) %>%
  summarise(
    proportion = mean(vote_conservative, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n())
  )

# Create the plot
ggplot(conservative_votes_by_province, aes(x = reorder(province, -proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have provinces on the y-axis
  labs(x = "Proportion", y = "Province", title = "Estimated Proportion of Voting for Conservative by Province") +
  theme_minimal()

@@

conservative_votes_by_education <- cleaned_survey_data %>%
  group_by(education) %>%
  summarise(
    proportion = mean(vote_conservative, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n())
  ) %>%
  arrange(desc(proportion))  
ggplot(conservative_votes_by_education, aes(x = reorder(education, proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  
  labs(x = "Proportion", y = "Education", title = "Proportion of Votes for Conservative by Education Level") +
  theme_minimal()

>>>>

step_cons <- glm(vote_conservative ~ age+province+education+religion+income_before_tax,
          data=cleaned_survey_data, family = "binomial")
step_cons_counts <- cleaned_survey_data %>%
  group_by(age, province, education, religion, income_before_tax)%>%
  summarise(N_j=n(), .groups = 'keep')

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
Party_Name <- c("Conservative", "Liberal")
yPS_primary <- c(cons_estimate,lib_estimate)
yPS_AIC <- c(cons_aic ,lib_aic)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(ggplot2)
library(dplyr)
library(broom)

required_vars <- c("vote_liberal", "age", "sex", "province", "education", "religion", "income_before_tax")
missing_vars <- setdiff(required_vars, names(cleaned_survey_data))

if(length(missing_vars) > 0) {
  stop(paste("The following required variables are missing in the data frame:", paste(missing_vars, collapse=", ")))
}
model_lib <- lm(vote_liberal ~ age + sex + province + education + religion + income_before_tax, data=cleaned_survey_data)
model_summary_lib <- broom::tidy(model_lib, conf.int = TRUE)

ggplot(model_summary_lib, aes(x = estimate, y = reorder(term, estimate))) +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Estimate", y = "Coefficient", title = "Coefficient Estimates for Liberal Vote Prediction") +
  theme_minimal()

@@

@

# Assuming 'cleaned_survey_data' contains 'province' and 'vote_liberal',
# where 'vote_liberal' is a binary (0/1) indicating whether a vote for liberal was cast.

# Calculate proportions and confidence intervals by province for the liberal party
liberal_votes_by_province <- cleaned_survey_data %>%
  group_by(province) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  # Drop grouping for further manipulation if needed
  ) %>%
  ungroup()  # Ensure we remove the grouping for the ggplot

# Create the plot for the liberal party
ggplot(liberal_votes_by_province, aes(x = reorder(province, -proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have provinces on the y-axis
  labs(x = "Proportion", y = "Province", title = "Estimated Proportion of Voting for Liberal by Province") +
  theme_minimal()

@@

# Assuming 'cleaned_survey_data' is your dataframe and it contains the variables 'education' and 'vote_liberal'
# where 'vote_liberal' is a binary (0/1) indicating whether a vote for liberal was cast.

liberal_votes_by_education <- cleaned_survey_data %>%
  group_by(education) %>%
  summarise(
    proportion = mean(vote_liberal, na.rm = TRUE),
    lower = proportion - qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    upper = proportion + qnorm(0.975) * sqrt(proportion * (1 - proportion) / n()),
    .groups = 'drop'  # This will automatically ungroup the data (requires dplyr version 1.0.0 or higher)
  ) %>%
  arrange(desc(proportion))

ggplot(liberal_votes_by_education, aes(x = reorder(education, proportion), y = proportion)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +  # Flip the axes to have education levels on the y-axis
  labs(x = "Proportion", y = "Education", title = "Proportion of Votes for Liberal by Education Level") +
  theme_minimal()
